---
title: " Workshop 2 portfolio analytics"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: ""
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse) # loads dplyr, ggplot2, and others
library(tidyquant)
library(fpp2)
library(ggpmisc)
library(tsfe)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
```

```{r set-up-data,echo=FALSE}
tickers<-c("GRG","BT.A","ULVR")
data("ftse350")
ftse350 %>% filter(ticker %in% tickers) -> port
port_ret <- port %>%
  filter(variable=="Price") %>%
  arrange(ticker,date) %>% # ensure data is order by stock and then chronologically 
  group_by(ticker) %>% # this ensures the lag operating only operate with each symbol
  mutate(log_return=log(value)-log(lag(value)))

port_ret_ew <- port_ret %>%
  group_by(date) %>%
  summarise(ret=mean(log_return))

port_w <- port %>% 
  select(ticker,date,variable,value) %>%
  spread(variable,value)

port_ret_vw <-port_w %>%
  group_by(date) %>%
  mutate(weight=`Market Value`/sum(`Market Value`))

port_ret_vw <-port_ret_vw %>% 
  arrange(ticker,date) %>% 
  group_by(ticker) %>% 
  mutate(log_return=log(Price)-log(lag(Price))) %>%
  group_by(date) %>%
  summarise(ret=sum(weight*log_return))

port_ret_both<- port_ret_vw %>% rename(vw_ret=ret) %>%
  left_join(port_ret_ew %>% rename(ew_ret=ret),by="date")


```


## Welcome

In this tutorial, you will learn how to summarise a table of data, including:

* How to rank financial data in terms of size
* How filter stocks `filter()` 
* Calculating and visualisng log returns

### Setup

I've preloaded the packages for this tutorial with 

```{r eval = FALSE}
library(tidyverse) # loads dplyr, ggplot2, and others
library(tidyquant)
library(ggpmisc)
library(fpp2)
```

## Topic 1 data manipulation

### Exercise explore the FTSE350 data

> In the `tsfe` pacakge open the  `ftse350` data and using the glimpse function to explore the data.  Describe what you see?  Click `Run Code` to see the data


```{r summarise, exercise = TRUE, exercise.eval = FALSE}
tsfe::ftse350
ftse350 %>% glimpse
```

```{r summarise-check}
"Good job!"
```


> What is the top 100 stocks in terms of market capitalisation? Challenge: filter the data to rank stocks on market size (Hint: some code to start `ftse350 %>% filter(variable=="Market Value") %>% group_by(date) %>%`)
    
```{r filter, exercise = TRUE}
    
```

```{r filter-solution}
ftse350 %>%
  filter(variable=="Market Value") %>%  
  group_by(date) %>% 
  mutate(rank = min_rank(desc(value))) %>% 
  filter(rank %in% c(1:100))
```
    
<div id="filter-hint">
**Hint:** Use`min_rank(desc(avg_delay))` to rank `avg_delay` (for example) such that the largest delay receives rank one. 
</div>

```{r filter-check}
"Great work!"
```


### Exercise extract the prices and market values for the following three stocks. Name the resultant dataframe `port`

1. Greggs 
2. BT
3. Unilever


```{r filtering-by-name, exercise=TRUE}
tickers<-c("GRG","BT.A","ULVR") ## these are the tickers of the above stocks
```

```{r filtering-by-name-solution}
tickers<-c("GRG","BT-A","ULVR")
ftse350 %>% filter(ticker %in% tickers) -> port
```

## Topic 2 Calculating and visualising returns

The port data which you created is in `tidy` form, which is one observation per row.  For the purposes of portfolio analytics we will need to filter on one variable to create returns.

### Exercise log returns
Create log returns for each daily price series using the `mutate` function in the `dplyr` package. Recall the formula

$$r_t=ln(P_t)-ln(P_{t-1})$$

**Hint:** use the `lag()` in the `mutate` to call $P_{t-1}$

```{r log-returns, exercise=TRUE}

```

```{r log-returns-solution, exercise=TRUE}

port_ret <- port %>%
  filter(variable=="Price") %>%
  arrange(ticker,date) %>% # ensure data is order by stock and then chronologically 
  group_by(ticker) %>% # this ensures the lag operating only operate with each symbol
  mutate(log_return=log(value)-log(lag(value)))

port_ret %>% head() # A sanity check to ensure that log returns are calculated correctly
```


### Exercise how to plot returns for comparison 1
Plot each log returns series in such a way as to compare their volatility over time using the aesthetic `colour`. Hint: try using `facet_wrap` in `ggplot2` to create separate plots.

```{r plot-price, excercise=T}

```

```{r plot-price-solution}
port_ret %>% 
  ggplot(aes(x=date,y=log_return,colour=ticker)) + 
  geom_line()
```

### Exercise how to plot returns for comparison 2
Plot each log returns series in such a way as to compare their volatility over time. Hint: try using `facet_wrap` in `ggplot2` to create separate plots.

```{r plot-price1}

```

```{r plot-price1-solution}
port_ret %>% 
  ggplot(aes(x=date,y=log_return,colour=ticker)) + 
  geom_line() +
  facet_wrap(~ticker)
```

Another way to compare these series is to consider the extreme (or outlying) returns.  Conventionally, you might want to consider values greater than 95^th^ percentile by using `stat_peak()` and `stat_valley()` from the `ggpmisc` package.  The latter package may have to be installed on your machine.

```{r peaks and troughs plot}
# install.packages("ggpmisc")
library(ggpmisc)
 port_ret %>% 
  ggplot(aes(x=date,y=log_return,colour=ticker)) + 
  geom_line() +
   stat_peaks(colour = "red",ignore_threshold = 0.95) +
   stat_valleys(colour = "blue",ignore_threshold =0.95) +
   facet_wrap(~ticker)
```

From the plot we can see that Greggs has 2 days where the returns are above the 95th percentile of peak, while BT and ULVR have one.Furthermore, Greggs and ulvr have 2 days where returns are below the 95th percentile of trough.  When doing your project how would to deal with outlying observations?

> In our experience, the unexpected is usually not an "outlier", or an aberrant point by rather a systematic pattern in some part of the data - Gelman et al. (2020) 


## Topic 3 Portfolio analytics

This topic you will calculate two daily portfolio return series for a portfolio containing the three stocks using log returns and market value.

### Exercise 1 equally weighted portfolio returns 

Create an equally weighted returns series for the three stocks, then plot the resultant return series
Hint: the mean is a equally weighted statistic.

Then using the portfolio return formula

$$r_{p,t} \approx \sum_{i=1}^{N}w_ir_{it}$$


```{r eq-weight-returns, exercise=TRUE}

```

```{r eq-weight-returns-solution}
port_ret_ew <- port_ret %>%
  group_by(date) %>%
  summarise(ret=mean(log_return))

port_ret_ew %>%
    ggplot(aes(x=date,y=ret)) +
  geom_line() 
```

### Exercise calculate value weighted

Create an value weighted returns series for the three stocks, then plot the resultant return series
Hint: use the data with both price and market value

$$w_{it}= \frac{V_{it}}{\sum V_{it}} \text{ where } V_{it}=Quantity \times P_{it}$$

Then using the portfolio return formula

$$r_{p,t} \approx \sum_{i=1}^{N}w_ir_{it}$$


```{r reshape the data}
port_w <- port %>% 
  select(ticker,date,variable,value) %>%
  spread(variable,value)
```


```{r vw-weight-returns, exercise=TRUE}

```


```{r vw-weight-returns-solution}
port_ret_vw <-port_w %>%
  group_by(date) %>%
  mutate(weight=`Market Value`/sum(`Market Value`))

# Sanity check
port_ret_vw %>% 
  summarise(tw=sum(weight)) %>%
  filter(tw>1.0001)

# market value weighted returns
port_ret_vw <-port_ret_vw %>% 
  arrange(ticker,date) %>% 
  group_by(ticker) %>% 
  mutate(log_return=log(Price)-log(lag(Price))) %>%
  group_by(date) %>%
  summarise(ret=sum(weight*log_return))
```

### Exercise merge returns

Combine the portfolio returns using `left_join()` . Hint: you need to choose a merging variable which unique identifies the portfolio returns time series.  Name the new object `port_ret_both`

```{r combine, exercise}

```


```{r combine-solution}
port_ret_both<- port_ret_vw %>% rename(vw_ret=ret) %>%
  left_join(port_ret_ew %>% rename(ew_ret=ret),by="date")
```

### Exercise plot returns and describe
Plot and visually compare the value-weight returns to the equally-weighted returns.  Provide some rationale for the differences? Hint: use `cumsum` to plot the wealth creation in the daily returns series

```{r compare_weightings}
```


```{r compare_weightings-solution}
port_ret_both %>%
  ungroup() %>%
  drop_na() %>%
  mutate(wealth_vw=cumsum(vw_ret),
         wealth_ew=cumsum(ew_ret)) %>%
  select(date,wealth_vw,wealth_ew) %>%
  gather(port,wealth,-date) %>%
  ggplot(aes(x=date,y=wealth,colour=port)) + 
  geom_line()
```

## Topic 4 Test your knowledge

```{r logicals, echo = FALSE}
question("Explain when it is **not** appropriate to use the above approximation for portfolio returns",
         answer("When the returns are far from zero, for example when the time interval is long, for instance quarterly or annual returns which are likely to be far from zero", correct = TRUE),
         answer("When the returns are close to zero, for example when the time interval is long, for instance quarterly or annual returns which are likely to be far from zero."),
         answer("When the returns are far from zero, for example when the time interval is short, for instance daily."),
         answer("When the returns are far from 1"),
         allow_retry = TRUE
)
```



```{r logicals1, echo = FALSE}
question("Why would an analyst prefer to use a total returns rather than price returns when asessing the performance of a portfolio investment?",
         answer("Total returns capture the income gains from holding an stock"),
         answer("Total returns capture the income and capital gains from holding an stock",correct = TRUE),
         answer("Total returns capture the capital and price gains from holding an stock"),
         answer("Total returns capture capital income and CEO turnover activity"),
         allow_retry = TRUE
)
```



```{r logicals2, echo = FALSE}
question("What investment assumption does the total return calculation make?",
         answer("In the long term dividend income becomes an increasingly important part of the total return opportunity for investing.", correct = TRUE),
         answer("In the long term dividend income becomes an decreasingly important part of the total return opportunity for investing."),
         answer("In the long term dividend income is not an important part of the total return opportunity for investing."),
         answer("Total returns capture capital income and merger activity"),
         allow_retry = TRUE
)
```

